---
title: "Case Study 1"
format: html
editor: visual
---

## Setup and Data Import

```{r}
library(tidyverse)
# pull in the data
world_castles <- read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-09-01/world_castles.csv"
)
```

```{r}
glimpse(world_castles)
```

```{r}
dim(world_castles)

summary(world_castles)

colSums(is.na(world_castles))
```

```{r}
world_castles |>
  count(category)

world_castles |>
  count(country, sort = TRUE)
```

## Data Cleaning

```{r}
# Create a cleaned version of the original castles data
castles_clean <- world_castles |>
  
  # Remove the 5 observations without an ISO country code
  filter(!is.na(iso)) |>
  
  # Remove the one observation with missing pageview information
  filter(!is.na(pageviews))
```

```{r}
# Add an approximate age variable for landmarks with a known founding year
# Since the data are from 2026, age is calculated as 2026 minus founding year
castles_clean <- castles_clean |>
  mutate(age = 2026 - year)
```

## Initial Data Exploration

### Landmark Categories

```{r}
# Count the number of landmarks in each category and display the results
# as a bar chart
castles_clean |>
  count(category) |>
  ggplot(aes(x = reorder(category, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Number of Landmarks by Category",
    x = "Landmark Category",
    y = "Number of Landmarks"
  ) +
  theme_minimal()
```

### Distribution of Wikipedia Pageviews

```{r}
# Examine the distribution of annual Wikipedia pageviews
# A histogram shows whether most landmarks have similar popularity
# or whether a small number receive substantially more views
ggplot(castles_clean, aes(x = pageviews)) +
  geom_histogram(bins = 50) +
  labs(
    title = "Distribution of Wikipedia Pageviews",
    x = "Annual Wikipedia Pageviews",
    y = "Number of Landmarks"
  ) +
  theme_minimal()
```

```{r}
# Pageviews are highly right-skewed, so a log10 scale makes the
# differences among less-famous landmarks easier to see
ggplot(castles_clean, aes(x = pageviews)) +
  geom_histogram(bins = 50) +
  scale_x_log10() +
  labs(
    title = "Distribution of Wikipedia Pageviews",
    subtitle = "Pageviews shown on a log10 scale",
    x = "Annual Wikipedia Pageviews (log scale)",
    y = "Number of Landmarks"
  ) +
  theme_minimal()
```

```{r}
# Identify the 15 landmarks with the most Wikipedia pageviews
castles_clean |>
  arrange(desc(pageviews)) |>
  select(name, country, category, year, pageviews, sitelinks, fame_rank) |>
  slice_head(n = 15)
```

### Distribution of Founding Years

```{r}
# Examine when landmarks in the dataset were founded
# Rows with unknown founding years are excluded only for this analysis
castles_clean |>
  filter(!is.na(year)) |>
  ggplot(aes(x = year)) +
  geom_histogram(binwidth = 100) +
  labs(
    title = "Distribution of Landmark Founding Years",
    x = "Founding Year",
    y = "Number of Landmarks"
  ) +
  theme_minimal()
```

## Landmark Popularity

### Popularity by Landmark Category

```{r}
# Compare Wikipedia popularity across the four landmark categories
# A log10 scale is used because pageviews are highly right-skewed
ggplot(castles_clean, aes(x = category, y = pageviews)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(
    title = "Wikipedia Pageviews by Landmark Category",
    x = "Landmark Category",
    y = "Annual Wikipedia Pageviews (log scale)"
  ) +
  theme_minimal()
```

```{r}
# Calculate summary statistics for pageviews within each landmark category

castles_clean |>
  group_by(category) |>
  summarise(
    n = n(),
    median_pageviews = median(pageviews),
    mean_pageviews = mean(pageviews),
    max_pageviews = max(pageviews)
  ) |>
  arrange(desc(median_pageviews))
```

### Founding Year and Popularity

```{r}
# Explore the relationship between founding year and Wikipedia popularity
# Pageviews are displayed on a log10 scale because of their strong skew
castles_clean |>
  filter(!is.na(year)) |>
  ggplot(aes(x = year, y = pageviews)) +
  geom_point(alpha = 0.3) +
  scale_y_log10() +
  labs(
    title = "Landmark Founding Year vs. Wikipedia Pageviews",
    x = "Founding Year",
    y = "Annual Wikipedia Pageviews (log scale)"
  ) +
  theme_minimal()
```

### Wikipedia Sitelinks and Pageviews

```{r}
# Compare two measures of landmark popularity:
# annual Wikipedia pageviews and the number of Wikipedia language editions
# containing an article about the landmark
ggplot(castles_clean, aes(x = sitelinks, y = pageviews)) +
  geom_point(alpha = 0.3) +
  scale_y_log10() +
  labs(
    title = "Wikipedia Language Coverage vs. Pageviews",
    x = "Number of Wikipedia Sitelinks",
    y = "Annual Wikipedia Pageviews (log scale)"
  ) +
  theme_minimal()
```

```{r}
# Calculate the correlation between sitelinks and pageviews
# This summarizes the strength of the linear relationship between
# the two measures of popularity
cor(castles_clean$sitelinks, castles_clean$pageviews)
```

```{r}
# Recalculate the correlation using log-transformed pageviews
# This reduces the influence of a small number of extremely
# high-pageview landmarks
cor(castles_clean$sitelinks, log10(castles_clean$pageviews + 1))
```

## Geographic Exploration

### Popularity by Country

```{r}
# Compare the typical popularity of landmarks across countries
# Only countries with at least 30 landmarks are included so that countries
# represented by only a few unusually famous landmarks do not dominate the results
country_popularity <- castles_clean |>
  group_by(country) |>
  summarise(
    n = n(),
    median_pageviews = median(pageviews),
    mean_pageviews = mean(pageviews)
  ) |>
  filter(n >= 30) |>
  arrange(desc(median_pageviews))

# Display the 15 countries with the highest median pageviews
country_popularity |>
  slice_head(n = 15)
```

```{r}
# Plot the 15 countries with the highest median landmark pageviews
country_popularity |>
  slice_head(n = 15) |>
  ggplot(aes(
    x = reorder(country, median_pageviews),
    y = median_pageviews
  )) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Countries with the Most Popular Landmarks",
    subtitle = "Countries with at least 30 landmarks in the dataset",
    x = "Country",
    y = "Median Annual Wikipedia Pageviews"
  ) +
  theme_minimal()
```

### Landmark Types Across Countries

```{r}
# Identify the 10 countries containing the most landmarks in the dataset
top_countries <- castles_clean |>
  count(country, sort = TRUE) |>
  slice_head(n = 10) |>
  pull(country)

# For each of the top 10 countries, calculate the proportion of landmarks
# belonging to each category
country_category <- castles_clean |>
  filter(country %in% top_countries) |>
  count(country, category) |>
  group_by(country) |>
  mutate(proportion = n / sum(n)) |>
  ungroup()

# Display the category composition of landmarks in the top 10 countries
ggplot(
  country_category,
  aes(x = reorder(country, proportion), y = proportion, fill = category)
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Landmark Types in the Countries with the Most Landmarks",
    x = "Country",
    y = "Proportion of Landmarks",
    fill = "Landmark Category"
  ) +
  theme_minimal()
```

```{r}
# Calculate median pageviews for each landmark category within the
# 10 countries containing the most landmarks in the dataset
country_category_popularity <- castles_clean |>
  filter(country %in% top_countries) |>
  group_by(country, category) |>
  summarise(
    n = n(),
    median_pageviews = median(pageviews),
    .groups = "drop"
  )

# Display the results 
country_category_popularity
```

```{r}
# Compare median Wikipedia pageviews across landmark categories
# within each of the 10 countries containing the most landmarks
ggplot(
  country_category_popularity,
  aes(
    x = category,
    y = median_pageviews,
    fill = category
  )
) +
  geom_col() +
  facet_wrap(~ country) +
  scale_y_log10() +
  labs(
    title = "Landmark Popularity by Category and Country",
    subtitle = "Ten countries with the most landmarks in the dataset",
    x = "Landmark Category",
    y = "Median Annual Wikipedia Pageviews (log scale)",
    fill = "Landmark Category"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Landmark History

### Founding Years by Landmark Category

```{r}
# Compare the distribution of founding years across landmark categories
castles_clean |>
  filter(!is.na(year)) |>
  ggplot(aes(x = category, y = year)) +
  geom_boxplot() +
  labs(
    title = "Founding Years by Landmark Category",
    x = "Landmark Category",
    y = "Founding Year"
  ) +
  theme_minimal()
```

```{r}
# Summarize founding years for each landmark category
# Median is useful because a small number of extremely ancient landmarks
# can strongly affect the mean founding year
castles_clean |>
  filter(!is.na(year)) |>
  group_by(category) |>
  summarise(
    n_with_year = n(),
    median_year = median(year),
    mean_year = mean(year),
    earliest_year = min(year),
    latest_year = max(year)
  ) |>
  arrange(median_year)
```

## Additional Popularity Exploration

### Comparing Measures of Fame

```{r}
# Identify the landmarks appearing in the largest number of
# Wikipedia language editions
castles_clean |>
  arrange(desc(sitelinks)) |>
  select(
    name, country, category, year,
    sitelinks, pageviews, fame_rank
  ) |>
  slice_head(n = 15)
```

```{r}
# Identify the landmarks receiving the most annual Wikipedia pageviews
# This can be compared with the sitelinks ranking to see whether
# international coverage and page traffic identify the same landmarks
castles_clean |>
  arrange(desc(pageviews)) |>
  select(
    name, country, category, year,
    sitelinks, pageviews, fame_rank
  ) |>
  slice_head(n = 15)
```

```{r}
# Calculate each landmark's percentile rank for pageviews and sitelinks
# Values near 1 indicate that a landmark ranks very highly compared
# with the rest of the dataset
castles_fame <- castles_clean |>
  mutate(
    pageview_percentile = percent_rank(pageviews),
    sitelink_percentile = percent_rank(sitelinks)
  )
```

```{r}
# Find landmarks in the top 10% for Wikipedia language coverage
# but below the median for annual pageviews
# These may be internationally documented landmarks that receive
# relatively modest current Wikipedia traffic
castles_fame |>
  filter(
    sitelink_percentile >= 0.90,
    pageview_percentile < 0.50
  ) |>
  arrange(desc(sitelinks)) |>
  select(
    name, country, category,
    sitelinks, pageviews,
    sitelink_percentile, pageview_percentile
  ) |>
  slice_head(n = 15)
```

```{r}
# Find landmarks in the top 10% for annual pageviews
castles_fame |>
  filter(
    pageview_percentile >= 0.90,
    sitelink_percentile < 0.50
  ) |>
  arrange(desc(pageviews)) |>
  select(
    name, country, category,
    sitelinks, pageviews,
    sitelink_percentile, pageview_percentile
  ) |>
  slice_head(n = 15)

```

## Multivariate Exploration

### Founding Year and Popularity by Category

```{r}
# Explore the relationship between founding year and Wikipedia popularity
# separately for castles, fortresses, palaces, and ruins

castles_clean |>
  filter(!is.na(year)) |>
  ggplot(aes(x = year, y = pageviews)) +
  geom_point(alpha = 0.3) +
  scale_y_log10() +
  facet_wrap(~ category) +
  labs(
    title = "Founding Year and Popularity by Landmark Category",
    subtitle = "Only landmarks with known founding years are included",
    x = "Founding Year",
    y = "Annual Wikipedia Pageviews (log scale)"
  ) +
  theme_minimal()
```

```{r}
# Calculate the correlation between founding year and log-transformed
# pageviews separately for each landmark category
#
# The log transformation reduces the influence of a small number of
# landmarks with extremely high pageview counts
castles_clean |>
  filter(!is.na(year)) |>
  group_by(category) |>
  summarise(
    n = n(),
    correlation = cor(year, log10(pageviews + 1))
  )
```

### Century Distribution

```{r}
# Examine the existing century variable before using it
# This lets us understand how historical periods are coded in the dataset
castles_clean |>
  count(century, sort = TRUE)
```

## Geographic Maps

### Overall Geographic Distribution

```{r}
# Load world map boundaries included with ggplot2
world_map <- map_data("world")

# Plot the geographic locations of all landmarks in the dataset
# Each point represents one castle, fortress, palace, or ruin
ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "gray90",
    color = "white"
  ) +
  geom_point(
    data = castles_clean,
    aes(x = lon, y = lat),
    alpha = 0.4,
    size = 0.8
  ) +
  coord_quickmap() +
  labs(
    title = "Geographic Distribution of Landmarks",
    subtitle = "Each point represents one landmark in the dataset",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_minimal()
```

### Geographic Distribution by Category

```{r}
# Map landmarks by category to explore whether castles, fortresses,
# palaces, and ruins have different geographic distributions
ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "gray90",
    color = "white"
  ) +
  geom_point(
    data = castles_clean,
    aes(x = lon, y = lat, color = category),
    alpha = 0.6,
    size = 1
  ) +
  coord_quickmap() +
  labs(
    title = "Geographic Distribution by Landmark Category",
    subtitle = "Castles, fortresses, palaces, and ruins around the world",
    x = "Longitude",
    y = "Latitude",
    color = "Landmark Category"
  ) +
  theme_minimal()
```

```{r}
# Map the geographic distribution of each landmark category separately
# Faceting reduces overlap and makes it easier to compare the locations
# of castles, fortresses, palaces, and ruins
ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "gray90",
    color = "white"
  ) +
  geom_point(
    data = castles_clean,
    aes(x = lon, y = lat),
    alpha = 0.5,
    size = 0.7
  ) +
  facet_wrap(~ category, ncol = 2) +
  coord_quickmap() +
  labs(
    title = "Geographic Distribution by Landmark Category",
    subtitle = "Each panel shows one type of landmark",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_minimal()
```

### Geographic Distribution of Popular Landmarks

```{r}
# Divide landmarks into four popularity groups based on annual pageviews
# Percentile groups make geographic differences easier to visualize
# despite the extremely skewed pageview distribution
castles_geo <- castles_clean |>
  mutate(
    popularity_group = case_when(
      percent_rank(pageviews) >= 0.90 ~ "Top 10%",
      percent_rank(pageviews) >= 0.50 ~ "50th–90th percentile",
      TRUE ~ "Bottom 50%"
    )
  )
```

```{r}
# Map only landmarks in the top 10% of annual Wikipedia pageviews
# This allows us to explore where the dataset's most-viewed
# landmarks are geographically concentrated
ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "gray90",
    color = "white"
  ) +
  geom_point(
    data = castles_geo |>
      filter(popularity_group == "Top 10%"),
    aes(x = lon, y = lat),
    alpha = 0.7,
    size = 1
  ) +
  coord_quickmap() +
  labs(
    title = "Location of the Most Popular Landmarks",
    subtitle = "Landmarks in the top 10% of annual Wikipedia pageviews",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_minimal()
```

## Data Quality Exploration

### Approximate Founding Years

```{r}
# Calculate the proportion of approximate founding dates
# within each landmark category
castles_clean |>
  filter(!is.na(year)) |>
  group_by(category) |>
  summarise(
    n = n(),
    approximate_years = sum(year_approx == 1),
    proportion_approximate = mean(year_approx == 1)
  )
```

```{r}
# Compare the founding-year distributions for landmarks with
# exact versus approximate founding dates
castles_clean |>
  filter(!is.na(year)) |>
  mutate(
    year_type = if_else(
      year_approx == 1,
      "Approximate year",
      "Exact year"
    )
  ) |>
  ggplot(aes(x = year_type, y = year)) +
  geom_boxplot() +
  labs(
    title = "Exact vs. Approximate Landmark Founding Years",
    x = "Type of Founding Year",
    y = "Founding Year"
  ) +
  theme_minimal()
```

## Multivariate Exploration: Country, Category, and Popularity

```{r}
# Identify the five countries with the largest number of landmarks
# Limiting the analysis to five countries keeps the visualization readable
top_5_countries <- castles_clean |>
  count(country, sort = TRUE) |>
  slice_head(n = 5) |>
  pull(country)


# Calculate median pageviews for each landmark category within each country
# Category-country combinations with fewer than 20 landmarks are removed
# so that medians are not based on very small groups
country_category_comparison <- castles_clean |>
  filter(country %in% top_5_countries) |>
  group_by(country, category) |>
  summarise(
    n = n(),
    median_pageviews = median(pageviews),
    .groups = "drop"
  ) |>
  filter(n >= 20)


# Display the values before graphing them
country_category_comparison
```

```{r}
# Compare the typical popularity of landmark categories across countries
ggplot(
  country_category_comparison,
  aes(
    x = category,
    y = median_pageviews,
    group = country,
    color = country
  )
) +
  geom_point(size = 3) +
  geom_line() +
  scale_y_log10() +
  labs(
    title = "Landmark Category and Popularity Across Countries",
    subtitle = "Only category-country groups with at least 20 landmarks",
    x = "Landmark Category",
    y = "Median Annual Wikipedia Pageviews (log scale)",
    color = "Country"
  ) +
  theme_minimal()
```


### Zero Wikipedia Pageviews

A substantial number of landmarks have zero recorded Wikipedia pageviews.
Because zero values cannot be displayed on a standard log10 scale, it is useful
to examine whether zero pageviews occur equally across landmark categories.

```{r}
# Calculate the proportion of landmarks with zero Wikipedia pageviews
# within each landmark category
zero_pageviews_by_category <- castles_clean |>
  group_by(category) |>
  summarise(
    n = n(),
    zero_pageviews = sum(pageviews == 0),
    proportion_zero = mean(pageviews == 0),
    .groups = "drop"
  ) |>
  arrange(desc(proportion_zero))

zero_pageviews_by_category
```

```{r}
# Visualize the percentage of landmarks with zero Wikipedia pageviews
# within each category
zero_pageviews_by_category |>
  ggplot(aes(
    x = reorder(category, proportion_zero),
    y = proportion_zero
  )) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Landmarks with Zero Wikipedia Pageviews by Category",
    x = "Landmark Category",
    y = "Percentage with Zero Pageviews"
  ) +
  theme_minimal()
```

### Missing Founding Years

Founding year is missing for a substantial proportion of the landmarks in the
dataset. Before examining relationships involving founding year, it is useful
to determine whether missingness differs across landmark categories.

```{r}
# Calculate the amount of missing founding-year information
# within each landmark category
missing_year_by_category <- castles_clean |>
  group_by(category) |>
  summarise(
    n = n(),
    missing_year = sum(is.na(year)),
    proportion_missing = mean(is.na(year)),
    .groups = "drop"
  ) |>
  arrange(desc(proportion_missing))

missing_year_by_category
```

```{r}
# Visualize the proportion of missing founding years by category
missing_year_by_category |>
  ggplot(aes(
    x = reorder(category, proportion_missing),
    y = proportion_missing
  )) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Missing Founding Years by Landmark Category",
    x = "Landmark Category",
    y = "Percentage Missing Founding Year"
  ) +
  theme_minimal()
```


## Fame Rank and Wikipedia Popularity


### Fame Rank and Wikipedia Popularity

The dataset also contains a `fame_rank` variable, where smaller values represent
more highly ranked landmarks. We can compare this ranking with Wikipedia
pageviews to see how closely the two measures of prominence agree.

```{r}
# Compare fame rank with Wikipedia pageviews
# log10(pageviews + 1) retains landmarks with zero pageviews
ggplot(
  castles_clean,
  aes(
    x = fame_rank,
    y = log10(pageviews + 1)
  )
) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Fame Rank vs. Wikipedia Pageviews",
    x = "Fame Rank",
    y = "log10(Annual Wikipedia Pageviews + 1)"
  ) +
  theme_minimal()
```

```{r}
# Use Spearman correlation because fame_rank is itself a ranking
# and the popularity variables are strongly skewed
cor(
  castles_clean$fame_rank,
  castles_clean$pageviews,
  method = "spearman"
)

cor(
  castles_clean$fame_rank,
  castles_clean$sitelinks,
  method = "spearman"
)
```



## Distribution of Wikipedia Sitelinks

### Distribution of Wikipedia Sitelinks

The number of Wikipedia sitelinks measures how many language editions contain
an article about a landmark. Before comparing sitelinks with pageviews, we can
examine the distribution of this measure on its own.

```{r}
# Examine the distribution of Wikipedia sitelinks
ggplot(castles_clean, aes(x = sitelinks)) +
  geom_histogram(binwidth = 5) +
  labs(
    title = "Distribution of Wikipedia Sitelinks",
    x = "Number of Wikipedia Sitelinks",
    y = "Number of Landmarks"
  ) +
  theme_minimal()
```

```{r}
# Compare the distribution of sitelinks across landmark categories
ggplot(castles_clean, aes(x = category, y = sitelinks)) +
  geom_boxplot() +
  labs(
    title = "Wikipedia Sitelinks by Landmark Category",
    x = "Landmark Category",
    y = "Number of Wikipedia Sitelinks"
  ) +
  theme_minimal()
```



## Century and Landmark Popularity

### Landmark Popularity by Century

In addition to examining how many landmarks were founded in each historical
period, we can compare whether landmarks from some centuries tend to receive
more Wikipedia attention than others.

```{r}
# Compare median Wikipedia pageviews across centuries
# Only centuries represented by at least 30 landmarks are included
century_popularity <- castles_clean |>
  filter(!is.na(century)) |>
  group_by(century) |>
  summarise(
    n = n(),
    median_pageviews = median(pageviews),
    mean_pageviews = mean(pageviews),
    .groups = "drop"
  ) |>
  filter(n >= 30)

century_popularity |>
  arrange(desc(median_pageviews))
```

```{r}
# Display median pageviews for the most well-represented centuries
century_popularity |>
  ggplot(aes(
    x = reorder(century, median_pageviews),
    y = median_pageviews
  )) +
  geom_col() +
  coord_flip() +
  scale_y_log10() +
  labs(
    title = "Median Wikipedia Pageviews by Century",
    subtitle = "Only centuries with at least 30 landmarks are included",
    x = "Century",
    y = "Median Annual Wikipedia Pageviews (log scale)"
  ) +
  theme_minimal()
```



