---
title: "Case Study 2"
format: html
editor: visual
---

## Front Matter

### Libraries

### Data Set

```{r}
lego <- read.csv("case-study2-data.csv")
```

## Exploratory Data Analysis: Theme, Pieces, Minifigures, and Unique Pieces (Gavin)

### 1. Data Preparation

#### 1.1 Variable Types and Missing Values

```{r}
View(lego)
```

```{r}
## EDA — Theme, Pieces, Minifigures, and Unique Pieces

### Data Preparation

# Convert Amazon_Price from text (e.g., "$19.99") to a number
lego$Amazon_Price <- as.numeric(gsub("[$,]", "", lego$Amazon_Price))

# Check the structure of the variables we will analyze
str(lego[, c("Amazon_Price", "Theme", "Pieces",
             "Minifigures", "Unique_Pieces")])

# Count missing values in our variables
colSums(is.na(lego[, c("Amazon_Price", "Theme", "Pieces",
                       "Minifigures", "Unique_Pieces")]))
```

#### 1.2 Cleaning LEGO Theme Names

```{r}
# Save the original theme names in case we need them later
lego$Theme_Original <- lego$Theme

# Convert theme names to plain ASCII text
# This removes problematic special characters
lego$Theme <- iconv(
  lego$Theme,
  from = "latin1",
  to = "ASCII",
  sub = ""
)

# Remove any extra spaces
lego$Theme <- trimws(lego$Theme)

# Check the most common themes after cleaning
head(sort(table(lego$Theme), decreasing = TRUE), 15)
```

### 2. Theme Analysis

#### 2.1 Number of LEGO Sets by Theme

```{r}
# Count how many LEGO sets belong to each theme
theme_counts <- sort(table(lego$Theme), decreasing = TRUE)

# Display the 15 most common themes
head(theme_counts, 15)
```

#### 2.2 LEGO Theme Prevalence — Word Cloud

```{r}
library(wordcloud)

# Count LEGO sets by theme
theme_counts <- sort(table(lego$Theme), decreasing = TRUE)

# Create clean theme names for the word cloud
theme_names <- iconv(
  names(theme_counts),
  from = "latin1",
  to = "ASCII",
  sub = ""
)

# Create the word cloud
set.seed(470)

wordcloud(
  words = theme_names,
  freq = as.numeric(theme_counts),
  min.freq = 2,
  max.words = 30,
  random.order = FALSE,
  rot.per = 0,
  scale = c(3.5, 0.7),
  colors = c(
    "#1F77B4",  # Blue
    "#FF7F0E",  # Orange
    "#2CA02C",  # Green
    "#D62728",  # Red
    "#9467BD",  # Purple
    "#17BECF",  # Teal
    "#E377C2",  # Pink
    "#BCBD22",  # Olive
    "#8C564B",  # Brown
    "#E6A23C",  # Gold
    "#264653",  # Dark teal
    "#E76F51"   # Coral
  ),
  ordered.colors = FALSE
)
```

#### 2.3 Median Amazon Price by LEGO Theme

```{r}
library(dplyr)
library(ggplot2)

# Calculate median Amazon price and number of priced sets by theme
theme_prices <- lego |>
  filter(!is.na(Theme), !is.na(Amazon_Price)) |>
  group_by(Theme) |>
  summarise(
    n = n(),
    median_price = median(Amazon_Price),
    .groups = "drop"
  ) |>
  filter(n >= 10)

# Plot median Amazon price for themes with at least 10 priced sets
ggplot(theme_prices,
       aes(x = reorder(Theme, median_price), y = median_price)) +
  geom_col(fill = "#2878B5") +
  coord_flip() +
  scale_y_continuous(labels = scales::label_dollar()) +
  labs(
    title = "Median Amazon Price by LEGO Theme",
    subtitle = "Themes with at least 10 sets with available Amazon prices",
    x = "LEGO Theme",
    y = "Median Amazon Price"
  ) +
  theme_minimal(base_size = 12)
```

#### 2.4 LEGO Theme and Amazon Price Distribution — Heat Map

```{r}
library(dplyr)
library(ggplot2)

# Keep sets with known themes and Amazon prices
theme_heatmap <- lego |>
  filter(!is.na(Theme), !is.na(Amazon_Price)) |>

  # Group Amazon prices into ranges
  mutate(
    Price_Range = cut(
      Amazon_Price,
      breaks = c(0, 25, 50, 100, 200, Inf),
      labels = c("$0–25", "$25–50", "$50–100",
                 "$100–200", "$200+"),
      include.lowest = TRUE
    )
  ) |>

  # Count sets in each theme and price range
  count(Theme, Price_Range, name = "Count") |>

  # Keep themes with at least 10 sets with known prices
  group_by(Theme) |>
  filter(sum(Count) >= 10) |>
  ungroup()

# Create heat map
ggplot(
  theme_heatmap,
  aes(
    x = Price_Range,
    y = reorder(Theme, Count, FUN = sum),
    fill = Count
  )
) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_gradient(
    low = "#EAF3F8",
    high = "#145A86"
  ) +
  labs(
    title = "LEGO Theme Distribution Across Amazon Price Ranges",
    subtitle = "Number of sets in each theme and price category",
    x = "Amazon Price Range",
    y = "LEGO Theme",
    fill = "Number of Sets"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

### 2.5 Percentage Distribution of Amazon Prices Within Each LEGO Theme

```{r}
# Calculate the percentage of sets in each price range within each theme
theme_heatmap_pct <- theme_heatmap |>
  group_by(Theme) |>
  mutate(
    Percentage = 100 * Count / sum(Count)
  ) |>
  ungroup()

# Create percentage-based heat map
ggplot(
  theme_heatmap_pct,
  aes(
    x = Price_Range,
    y = reorder(Theme, Percentage, FUN = max),
    fill = Percentage
  )
) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_gradient(
    low = "#EAF3F8",
    high = "#145A86",
    limits = c(0, 100),
    labels = function(x) paste0(x, "%")
  ) +
  labs(
    title = "Amazon Price Distribution Within Each LEGO Theme",
    subtitle = "Percentage of priced sets in each price range",
    x = "Amazon Price Range",
    y = "LEGO Theme",
    fill = "Percentage"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

### 3. Pieces Analysis

#### 3.1 Distribution of LEGO Piece Counts

```{r}
# Summary statistics for the number of pieces
summary(lego$Pieces)

# Histogram showing the distribution of piece counts
ggplot(lego, aes(x = Pieces)) +
  geom_histogram(
    bins = 40,
    fill = "#2878B5",
    color = "white"
  ) +
  scale_x_continuous(
    labels = scales::label_comma()
  ) +
  labs(
    title = "Distribution of LEGO Set Piece Counts",
    subtitle = "Number of pieces per LEGO set",
    x = "Number of Pieces",
    y = "Number of LEGO Sets"
  ) +
  theme_minimal(base_size = 12)
```

#### 3.2 Relationship Between Pieces and Amazon Price

```{r}
# Keep sets with known piece counts and Amazon prices
pieces_price <- lego |>
  filter(!is.na(Pieces), !is.na(Amazon_Price))

# Calculate the Pearson correlation
cor(pieces_price$Pieces, pieces_price$Amazon_Price)

# Scatterplot of piece count and Amazon price
ggplot(pieces_price, aes(x = Pieces, y = Amazon_Price)) +
  geom_point(
    color = "#2878B5",
    alpha = 0.5,
    size = 2
  ) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    color = "#E76F51",
    linewidth = 1
  ) +
  scale_x_continuous(labels = scales::label_comma()) +
  scale_y_continuous(labels = scales::label_dollar()) +
  labs(
    title = "Relationship Between LEGO Pieces and Amazon Price",
    subtitle = "Scatterplot with fitted linear trend",
    x = "Number of Pieces",
    y = "Amazon Price"
  ) +
  theme_minimal(base_size = 12)
```

#### 3.3 Log-Log Relationship Between Pieces and Amazon Price

```{r}
# Keep observations with positive piece counts and Amazon prices
pieces_price_log <- pieces_price |>
  filter(Pieces > 0, Amazon_Price > 0)

# Create a scatterplot with logarithmic axes
ggplot(
  pieces_price_log,
  aes(x = Pieces, y = Amazon_Price)
) +
  geom_point(
    color = "#2878B5",
    alpha = 0.5,
    size = 2
  ) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    color = "#E76F51",
    linewidth = 1
  ) +
  scale_x_log10(
    labels = scales::label_comma()
  ) +
  scale_y_log10(
    labels = scales::label_dollar()
  ) +
  labs(
    title = "LEGO Pieces vs. Amazon Price (Log-Log Scale)",
    subtitle = "Logarithmic axes highlight patterns across set sizes",
    x = "Number of Pieces (Log Scale)",
    y = "Amazon Price (Log Scale)"
  ) +
  theme_minimal(base_size = 12)
```

### 4. Minifigures Analysis

#### 4.1 Distribution of LEGO Minifigure Counts

```{r}
# Summary statistics for the number of minifigures
summary(lego$Minifigures)

# Count how many sets have each recorded minifigure count
table(lego$Minifigures, useNA = "ifany")

# Create a histogram of minifigure counts
ggplot(
  lego |> filter(!is.na(Minifigures)),
  aes(x = Minifigures)
) +
  geom_histogram(
    binwidth = 1,
    boundary = -0.5,
    fill = "#2878B5",
    color = "white"
  ) +
  scale_x_continuous(breaks = scales::breaks_pretty()) +
  labs(
    title = "Distribution of LEGO Minifigure Counts",
    subtitle = "Sets with recorded minifigure counts",
    x = "Number of Minifigures",
    y = "Number of LEGO Sets"
  ) +
  theme_minimal(base_size = 12)
```

#### 4.2 Relationship Between Minifigures and Amazon Price

```{r}
# Keep sets with known minifigure counts and Amazon prices
minifig_price <- lego |>
  filter(!is.na(Minifigures), !is.na(Amazon_Price))

# Calculate Pearson correlation
cor(minifig_price$Minifigures, minifig_price$Amazon_Price)

# Create a scatterplot with slight horizontal jitter
set.seed(470)

ggplot(minifig_price, aes(x = Minifigures, y = Amazon_Price)) +
  geom_jitter(
    width = 0.2,
    height = 0,
    color = "#2878B5",
    alpha = 0.5,
    size = 2
  ) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    color = "#E76F51",
    linewidth = 1
  ) +
  scale_x_continuous(breaks = scales::breaks_pretty()) +
  scale_y_continuous(labels = scales::label_dollar()) +
  labs(
    title = "Relationship Between LEGO Minifigures and Amazon Price",
    subtitle = "Recorded minifigure counts with fitted linear trend",
    x = "Number of Minifigures",
    y = "Amazon Price"
  ) +
  theme_minimal(base_size = 12)
```

### 5. Unique Pieces Analysis

#### 5.1 Distribution of Unique LEGO Piece Counts

```{r}
# Summary statistics for unique piece counts
summary(lego$Unique_Pieces)

# Create a histogram of unique piece counts
ggplot(
  lego |> filter(!is.na(Unique_Pieces)),
  aes(x = Unique_Pieces)
) +
  geom_histogram(
    bins = 40,
    fill = "#2878B5",
    color = "white"
  ) +
  scale_x_continuous(labels = scales::label_comma()) +
  labs(
    title = "Distribution of Unique LEGO Piece Counts",
    subtitle = "Number of distinct piece types per LEGO set",
    x = "Number of Unique Pieces",
    y = "Number of LEGO Sets"
  ) +
  theme_minimal(base_size = 12)
```

#### 5.2 Relationship Between Unique Pieces and Amazon Price

```{r}
# Keep sets with known unique piece counts and Amazon prices
unique_price <- lego |>
  filter(!is.na(Unique_Pieces), !is.na(Amazon_Price))

# Calculate Pearson correlation between unique pieces and Amazon price
cor(unique_price$Unique_Pieces, unique_price$Amazon_Price)

# Create a scatterplot with a fitted linear trend
ggplot(
  unique_price,
  aes(x = Unique_Pieces, y = Amazon_Price)
) +
  geom_point(
    color = "#2878B5",
    alpha = 0.5,
    size = 2
  ) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    color = "#E76F51",
    linewidth = 1
  ) +
  scale_x_continuous(labels = scales::label_comma()) +
  scale_y_continuous(labels = scales::label_dollar()) +
  labs(
    title = "Relationship Between Unique Pieces and Amazon Price",
    subtitle = "Scatterplot with fitted linear trend",
    x = "Number of Unique Pieces",
    y = "Amazon Price"
  ) +
  theme_minimal(base_size = 12)
```

#### 5.3 Relationship Between Total Pieces and Unique Pieces

```{r}
# Keep sets with both piece counts available
pieces_unique <- lego |>
  filter(!is.na(Pieces), !is.na(Unique_Pieces))

# Calculate the Pearson correlation
cor(pieces_unique$Pieces, pieces_unique$Unique_Pieces)

# Visualize the relationship
ggplot(
  pieces_unique,
  aes(x = Pieces, y = Unique_Pieces)
) +
  geom_point(
    color = "#2878B5",
    alpha = 0.5,
    size = 2
  ) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    color = "#E76F51",
    linewidth = 1
  ) +
  scale_x_continuous(labels = scales::label_comma()) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    title = "Relationship Between Total Pieces and Unique Pieces",
    subtitle = "Exploring the association between two potential price predictors",
    x = "Number of Pieces",
    y = "Number of Unique Pieces"
  ) +
  theme_minimal(base_size = 12)
```

#### 5.4 Log-Log Relationship Between Unique Pieces and Amazon Price

```{r}
# Keep observations with positive unique piece counts and Amazon prices
unique_price_log <- unique_price |>
  filter(Unique_Pieces > 0, Amazon_Price > 0)

# Create a log-log scatterplot
ggplot(
  unique_price_log,
  aes(x = Unique_Pieces, y = Amazon_Price)
) +
  geom_point(
    color = "#2878B5",
    alpha = 0.5,
    size = 2
  ) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    color = "#E76F51",
    linewidth = 1
  ) +
  scale_x_log10(
    labels = scales::label_comma()
  ) +
  scale_y_log10(
    labels = scales::label_dollar()
  ) +
  labs(
    title = "Unique LEGO Pieces vs. Amazon Price (Log-Log Scale)",
    subtitle = "Logarithmic axes highlight relationships across set sizes",
    x = "Number of Unique Pieces (Log Scale)",
    y = "Amazon Price (Log Scale)"
  ) +
  theme_minimal(base_size = 12)
```

### 6. Additional Data Quality Investigation

#### 6.1 Missing Minifigure Counts by LEGO Theme

```{r}
# Calculate missing minifigure counts by theme
minifig_missing <- lego |>
  filter(!is.na(Theme)) |>
  group_by(Theme) |>
  summarise(
    Total_Sets = n(),
    Missing_Minifigures = sum(is.na(Minifigures)),
    Missing_Percent = 100 * mean(is.na(Minifigures)),
    .groups = "drop"
  ) |>
  filter(Total_Sets >= 10) |>
  arrange(desc(Missing_Percent))

# Display missingness by theme
print(minifig_missing, n = Inf)

# Visualize missingness
ggplot(
  minifig_missing,
  aes(x = reorder(Theme, Missing_Percent), y = Missing_Percent)
) +
  geom_col(fill = "#2878B5") +
  coord_flip() +
  scale_y_continuous(
    labels = function(x) paste0(x, "%")
  ) +
  labs(
    title = "Missing Minifigure Data by LEGO Theme",
    subtitle = "Themes with at least 10 sets",
    x = "LEGO Theme",
    y = "Percentage Missing"
  ) +
  theme_minimal(base_size = 12)
```

### 7. Summary of Exploratory Findings

#### Theme

-   LEGO themes differ considerably in their typical Amazon prices. Creator Expert has the highest median price among the themes analyzed, while DOTS and Classic are among the lowest.

-   The word cloud shows that Star Wars, Friends, City, and NINJAGO are the most prevalent themes in the dataset.

-   The percentage-based heat map reveals that Amazon price distributions vary across themes, suggesting that Theme may be useful for predicting price.

#### Pieces

-   Total piece counts are strongly right-skewed, with a median of 198 pieces and a maximum of 6,020.

-   Pieces has a strong positive correlation with Amazon Price (r = 0.8117).

-   The log-log scatterplot shows an approximately linear positive relationship, suggesting that logarithmic transformations may be worth exploring during modeling.

#### Minifigures

-   Recorded minifigure counts are right-skewed, with a median of 2 and a maximum of 28.

-   Minifigures has a moderate positive correlation with Amazon Price (r = 0.5618).

-   Missing minifigure data is strongly associated with Theme. Several themes have 100% missing values, indicating that missingness may reflect differences in set characteristics or data collection.

#### Unique Pieces

-   Unique piece counts are right-skewed, with a median of 83 and a maximum of 1,067.

-   Unique_Pieces has a strong positive correlation with Amazon Price (r = 0.7061).

-   Pieces and Unique_Pieces are also strongly correlated (r = 0.7794), suggesting that they may contain overlapping predictive information.

#### Overall Conclusions

-   All three numerical variables examined have positive associations with Amazon Price, with Pieces showing the strongest pairwise linear correlation.

-   Theme also appears to be associated with price, although differences may partly reflect variation in set sizes.

-   Missing data, especially for Amazon_Price and Minifigures, will require careful consideration during predictive modeling.

-   Future modeling should evaluate whether Theme, Minifigures, and Unique_Pieces improve price predictions beyond what can be achieved using total piece count alone.
